\bf Model Visualization Techniques for a Social Network Model

### 2018-08-15 paper-child.Rnw
ugh what is happening with the fonts of the plots????

\noindent \textbf{Introduction to Network Structures:} Network data across fields have similar data structures. There are always units of observation and connections between those units. In a social network, the units of observations, called \textit{nodes} or \textit{actors}, may be people, while the connections, called \textit{edges} or \textit{ties}, are the relationships between people. Networks often change over time, like when new relationships are formed in a social network. The nodes and edges themselves can also have inherent variables of interest, e.g.\ the age, gender, and political affiliation of people in a social network and the type of relationship (friends, married, etc.). \\

@@

Because we require binary edges for SAOMs, we focus only on very strong collaborations, where $WPC$ is high. For a senate collaboration network $x$, edges are defined as
\begin{equation}\label{eq:wpc2}
  x_{ij} =
  \begin{cases}
                                   1 &  \text{if } WPC_{ij} > 0.25 \\
                                   0 & \text{if } WPC_{ij} \leq 0.25.
  \end{cases}
\end{equation}
The node-link diagram for the four senates during the Obama administration are shown in Figure~\ref{fig:senateplot}. We can see that Senate 111 is much more densely connected than the other three, which have one large, sparsely connected component with a few smaller connected groups throughout. Some prominent names are shown to demonstrate their place in the network.

\noindent \textbf{Models:} We first consider the 16 actor subset of the teenage friends and lifestyle data show in Figure~\ref{fig:alldataviz}. To this data, we fit three different SAOMs. Each SAOM used a simple rate function, $\alpha_m$, and an objective function with two or three parameters. The first model, M1, contains the absolute minimum number of parameters in the objective function $f_i(x)$: 
\begin{equation}\label{eq:m1}
f_i(x)^{M1} = \beta_1s_{i1} + \beta_2s_{i2},
\end{equation}
where $s_{i1}$ is the density network statistic and $s_{i2}$ is the reciprocity network statistic for actor $i$ at the current network state $x$. The second and third models, M2 and M3, contain one additional parameter each in the objective function which were determined by a Wald-type test provided in the \texttt{RSiena} software to be significant, with $p$-values less than 0.05 \citep{RSienaTest}. The M2 model contains an actor-level covariate parameter, and the M3 model contains an additional strutural effect in the objective function.
\begin{align}\label{eq:m2m3}
f_i(x)^{M2} & = \beta_1s_{i1} + \beta_2s_{i2} + \beta_3s_{i3} \\
f_i(x)^{M3} & = \beta_1s_{i1} + \beta_2s_{i2} + \beta_4s_{i4},
\end{align}

@@

set.seed(12345)
ggplot(data = jTT, aes(from_id = from, to_id = to)) +
  geom_net(aes(shape = as.factor(group)), directed = T, labelon = T,
           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15,
           colour = 'black', size=10, fontsize = 5,
           ecolour = c("grey40", "grey40","red", "grey40")) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none")
@
\caption{\label{fig:jtt}Realization of a JTT, where $i$ is the ego, $j$ is the alter, and $h$ is the intermediary. The covariate is represented by the shape of the nodes.}
\end{subfigure}\hfill
\begin{subfigure}[t]{.45\textwidth}
<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', out.width='.49\\textwidth'>>=
dade <- data.frame(from = c('i', 'i', 'h', 'j', 'i'), to = c('h', 'j', 'k', 'k','k'))
dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))

@@

set.seed(12345)
ggplot(data = dad, aes(from_id = from, to_id = to)) +
  geom_net(directed = T, labelon = T,
           ecolour = c("black", "black", "black","red", "black" ,"black" ),
           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, size=10, fontsize = 5) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none")
@
\caption{\label{fig:dad}Realization of a N22 between actors $i$ and $k$.}
\end{subfigure}
\caption{The additional network effects included in our models fit to the friends data. On the left, a JTT and on the right, a N22 between $i$ and $k$.}
\end{figure}

where $s_{i3} = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$, and $s_{i4} =  |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$. These statistics are known as the \emph{number of jumping transitive triplets} (JTT) and the \emph{number of doubly achieved distances two effect} (N22), respectively. The first statistic emphasizes triad relationships that are formed between actors from different covariate groups, while the other emphasizes indirect ties between actors. The covariate groups are determined by the student's drinking behavior. These additional effects are shown in Figure~\ref{fig:jtt} and Figure~\ref{fig:dad}.

@@

In this paper, we use three different model visualization techniques on SAOMs. First, Section~\ref{sec:minds} places the models in the data space. Then, Section~\ref{sec:collectmodels} explores visualizations of various collections of models. Finally, in Section~\ref{sec:algorithms} we use model visualization to explore the underlying algorithms in the SAOM fitting process. 
 %In terms of the senate data, then, the value of $\beta_3$ should dictate how willing a senator is to coauthor a bill on which the author of the bill has a different level of authorship, assuming there is an intermediary between the two who has the same authorship level as the first. %Comparison of the parameter estimates of these models fit to the two different data sets is given in Table~\ref{tab:meanests}.
%In addition, we fit two more models, M4 and M5 to the senate data that were found to be significant according to the Wald-type test:

@@

Every good data analysis includes both numerical and visual summaries of the data, so why restrict model description and diagnostics to numerical summaries? The concept of model visualization was developed to complement traditional model diagnostic tools such as $R^2$ values and residual plots \citep{modelvis}. When performing analyses, there are three different types of ``model": the model family, the model form, and the fitted model. The latter is primarily what one thinks of first when considering a model in a data analysis, where a specified model is fit to data, and parameter estimates are reported.% In the context of SAOMs, the \emph{fitted model} contains the form of the rate and objective functions, the estimated rate parameters, and the estimated objective function parameters. 
The \textit{model form} describes the the model \textit{before} the fitting process, defining which parameters are in the model within the context of the larger model family.% In SAOMs, the \emph{model form} includes description of the rate and objective functions and the variables therein that describe how the network evolves over time. 
Finally, the \textit{model family} is the broadest description of the model. This is the type of model that you wish to fit to the data, and is chosen based on the problem, data, and knowledge at hand. %For example, we chose to use a SAO \emph{model family} to model our network data over an exponential random graph model (ERGM) because we believe that actor-level variables effect network structure and formation, and we wanted to model the network changes over time.

The model family, the model form, and the fitted model can each be visualized according to the three principles of model visualization: we can view the model in the data space, visualize collections of models, and explore the process of fitting the model, not just the end result. Since we have already decided on our model family, SAOMs, we shift our focus to the fitted model and the model form. Specifically, we want to learn more about how the \emph{model form} we choose affects the \emph{fitted model} by using our example data sets and our visualization toolbox. We begin by introducing the five models that we fit to our example data. We then use the five models to guide our visual explorations of SAOMs under the three principles of model visualization. %We begin with novel tools and ideas to view a SAOM in the data space of a dynamic network. We then explore collections of the same models fit many times to the example data to see how the simulation processes in the \texttt{RSiena} ``black-box" affect the model fits. Finally, we look behind the scenes and into the individual steps of the continuous time Markov chain to learn more about how this ``hidden data" mechanism works and how it results in a fitted model.

The first way we hope to better understand SAOMs is by viewing the model(s) in the data space. In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for network models? We define the data space as the combined data structures of the network: the node, edge, and time information.
%We want to make inferences on parameters in the model, but the multi-level data structure of networks makes the ``region of inference" difficult to define.
% %For dynamic social networks, there are a few different data ``spaces":
% \begin{enumerate}
% \item the actors and their corresponding covariates,
% \item the edges and their correpsonding variables that describe the relationships between the nodes, and
% \item the time, both the continuous unobserved time and the discrete observed time points, over which the network evolves
% \end{enumerate}
%The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if there is an edge between two nodes, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side.

@@

The node, edge, and time data can be visualized together in various ways, and one tool that can bring the node, edge, and time data spaces together is the \texttt{R} package \texttt{geomnet} \citep{geomnet}. Different visual features in the node-link diagram can be tied to the underlying node or edge data. The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the lines between points can be used to represent the edge variables. %In a social network, node data might be age, gender, and/or occupation of the person in the network, and edge data might be length of connection between two people, how the people first met (school, work, church, etc.), and/or how often they interact. 
To view temporal changes, we can view the network at different timepoints side-by-side to see the evolution. Pulling all of this information together with \texttt{geomnet} allows the entire data space to be viewed at once.

@@

To demonstrate, we use \texttt{geomnet} to visualize the connections in the $111^{th}$ United States Senate at two different points: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate for 17 days, from January 3, 2009, to January 20, 2009, when she was in the middle of her second term as senator from New York. In that time, she authored two bills and was a cosponsor on 17 other bills. There are many Democratic senators who cosponsored both of her bills, giving their edges a $WPC$ of 1. So with Clinton included in the node-link diagram, the senate looks much more collaborative than it does without her. We then map the size of the node to the numer of bills authored by the senator, the shape to sex, color to party. Bills, sex, and party are all potential covariates in a SAOM. In addition, we visualize the strength of the tie by mapping the linewidth of the edge to the $WPC$ value between the two senators. In this single visualization, we have viewed node information, edge information, and time. We will use \texttt{geomnet} many ways to view the model in the data space.

@@

p1 <- ggplot(data = t1) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1$color,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Average: M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))

@@

For network data, we view an ``average" network through a sort of summary network drawn using the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1. To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then combine the 1,000 simulated instances of wave 2, and count up the number of times each edge appears in a network simulation. Then, we combine these 1,000 networks into a single network with edgeweight equal to the proportion of time that edge appears in our 1,000 simulations. This weighted edgelist is the network we draw in Figure~\ref{fig:summnetM1}. An edge is only drawn in the average network if it appears in at least 51 of the 1000 simulations, with edges that appear with greater frequency emphasized by thicker linewidths and a darker color. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data on the right of Figure~\ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

<<summary2, fig.height=2.5, fig.cap="Network differences between the average network from M1 and the first and second data waves. The green edges represent edges in the average network that are not in the data, while red edges are in the data but not in the average network. Black edges are in both the data and the average, and the edgeweight represents the rate at which that edge shows up in simulations.">>=
am_t1 <- as.matrix(igraph::as_adj(igraph::graph_from_data_frame(arrange(t1, parse_number(from), parse_number(to)))))[-17,-17]

diff1 <- am_t1 - fd2.w1
diff1[which(diff1==1)] <- 2
diff1[which(fd2.w1 == 1 & am_t1 == 1)]  <- 1
diff1 <- diff1 %>% dplyr::as_data_frame() 
names(diff1) <- paste0("V", 1:16)
diff1 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>%
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `2` = "InAvgNotW1",  `1`= "InBoth"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = "green")) %>%
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_1 
t1_1 <- arrange(t1_1, from, to)
p1 <- ggplot(data = t1_1) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_1$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Difference between\nWave 1 and Average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))

diff2 <- -am_t1 + fd2.w2 
diff2[which(diff2==1)] <- 2
diff2[which(fd2.w2 == 1 & am_t1 == 1)]  <- 1
diff2 <- dplyr::as_data_frame(diff2)
names(diff2) <- paste0("V", 1:16)
diff2 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>% 
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `-1` = "InAvgNotW2",  `1`= "InBoth", `2` = "InW2notAvg"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = "red", `-1` = "green")) %>% 
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_2 
t1_2$linewidth[is.na(t1_2$linewidth) & !is.na(t1_2$to)] <- 1
t1_2 <- arrange(t1_2, from, to)
t1_2 <- t1_2 %>% filter(!(from == 16 & is.na(to)))
p2 <- ggplot(data = t1_2) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_2$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Difference between\nWave 2 and Average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))
grid.arrange(p1, p2, nrow = 1)
@

In Figure~\ref{fig:summary2}, we show \textit{difference networks} between the average network from M1 and the first and second waves of data. The lack of model fit is very obvious here, with the green edges representing ties that only show up in the average network, and the red edges representing ties in the data that the model misses completely.  Again, the edgeweights demonstrate the frequency of the ties in the simulations from M1. On the left in Figure~\ref{fig:summary2}, we see the difference between the first wave of data and the average network. We only see green and black edges, meaning that the simulated networks consist primarily of edges in wave 1, plus some new edges that aren't in the first wave of data that are reciprocating the ties that are in the data. On the right in Figure~\ref{fig:summary2}, we see the difference between the second wave of data and the average network. Remember, the model is simulating the second wave of data, which is how the average network is created. There are six red edges, which are edges in wave 2 that do not appear at all in the average network. In addition, there are several green edges that are in the average network, but not in wave 2. Finally, there are a few black edges that appear in both. For all edges but the red ones, the linewidth represents the frequency of the edge in the simulations.

@@

We have varied model settings already by choosing models M1, M2, M3 to fit to our small friendship network data set. In Section~\ref{sec:samemodsamedat}, we fit these models to the data 1,000 times, and in this section, we explore simulations from these three models given the mean values of from the 1,000 fitted parameter values as the parameters in our models. From each of these three models using the means of parameter estimates as our fixed parameter values, we simulated 1,000 observations from each of the three models. From these simulations, we create visualizations that represents an average network using the same method described in Section~\ref{sec:minds}, shown in Figure~\ref{fig:averagenet} in the Appendix.

@@

>>>>

\noindent \textbf{Network Data Structure:} Network data across fields have similar data structures. There are always units of observation and connections between those units. In a social network, the units of observations, called \textit{nodes} or \textit{actors}, may be people, while the connections, called \textit{edges} or \textit{ties}, are the relationships between people. Networks often change over time, like when new relationships are formed in a social network. The nodes and edges themselves can also have inherent variables of interest, e.g.\ the age, gender, and political affiliation of people in a social network and the type of relationship (friends, married, etc.). \\

@@

Because we require binary edges for SAOMs, we focus only on very strong collaborations, where $WPC$ is high. For a senate collaboration network $x$, edges are defined as
\begin{equation}\label{eq:wpc2}
  x_{ij} =
  \begin{cases}
                                   1 &  \text{if } WPC_{ij} > 0.25 \\
                                   0 & \text{if } WPC_{ij} \leq 0.25.
  \end{cases}
\end{equation}
The node-link diagram for the four senates during the Obama administration are shown in Figure~\ref{fig:senateplot}. We can see that Senate 111 is much more densely connected than the other three, which have one large, sparsely connected component with a few smaller connected groups throughout. Some prominent names are shown to demonstrate their place in the network.  \\

\noindent \textbf{Models:} To our example data, we fit three different SAOMs. Each SAOM uses a simple rate function, $\alpha_m$, and an objective function with two or three parameters. The first model, M1, contains the absolute minimum number of parameters in the objective function $f_i(x)$: 
\begin{equation}\label{eq:m1}
f_i(x)^{M1} = \beta_1s_{i1}(x) + \beta_2s_{i2}(x),
\end{equation}
where $s_{i1}$ is the density network statistic and $s_{i2}$ is the reciprocity network statistic for actor $i$ at the current network state $x$ (see Table~\ref{tab:effects}. The second and third models, M2 and M3, contain one additional parameter each in the objective function which were determined by a Wald-type test provided in the \texttt{RSiena} software to be significant, with $p$-values less than 0.05 \citep{RSienaTest}. The M2 model contains an actor-level covariate parameter, and the M3 model contains an additional strutural effect in the objective function.
\begin{align}\label{eq:m2m3}
f_i(x,z)^{M2} & = \beta_1s_{i1}(x) + \beta_2s_{i2}(x) + \beta_3s_{i3}(x, z) \\
f_i(x)^{M3} & = \beta_1s_{i1}(x) + \beta_2s_{i2}(x) + \beta_4s_{i4}(x),
\end{align}

@@

set.seed(12345)
ggplot(data = jTT, aes(from_id = from, to_id = to)) +
  geom_net(aes(shape = as.factor(group)), directed = T, labelon = T,
           labelcolour='white',vjust = 0.5, hjust =0.5, arrowgap = .15,
           size=10, fontsize = 5, ecolour="black",
           linetype = c("solid", "solid","dotted", "solid")) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none") + 
  labs(title = expression(paste(s[i3], "(x,z)", "=",Sigma[j != h], x[ij], x[ih], x[hj] %.% 
                                I, "(", z[i], "=", z[h]!= z[j], ")")
                          )
       )
@
\caption{\label{fig:jtt}Realization of a JTT, where $i$ is the ego, $j$ is the alter, and $h$ is the intermediary. The covariate is represented by the shape of the nodes.}
\end{subfigure}\hfill
\begin{subfigure}[t]{.45\textwidth}
<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', out.width='.49\\textwidth'>>=
dade <- data.frame(from = c('i', 'i', 'h', 'k', 'i'), to = c('h', 'k', 'j', 'j','j'))
dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))

@@

set.seed(12345)
ggplot(data = dad, aes(from_id = from, to_id = to)) +
  geom_net(directed = T, labelon = T, ecolour="black",
           linetype = c("solid", "solid", "dotted", "solid","solid" ,"solid"),
           labelcolour='white',vjust = 0.5, hjust =0.5, arrowgap = .15, size=10, fontsize = 5) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none") + 
  labs(title = expression(paste(s[i4], "(x)", "=", "|{j:", x[ij],"=0, ", Sigma[h], x[ih], x[hj] >= 2, "}|" )
                          )
       )
@
\caption{\label{fig:dad}Realization of a N22 between actors $i$ and $j$. The dotted tie is encouraged if $\beta_4$ has a positive value.}
\end{subfigure}
\caption{The additional network effects included in our models fit to the friends data. On the left, a JTT between $i$ and $j$, who have different covariate values, and on the right, a N22 between $i$ and $j$.}
\end{figure}

\noindent where $s_{i3}(x,z) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$, and $s_{i4}(x) =  |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$. These statistics are known as the \emph{number of jumping transitive triplets} (JTT) and the \emph{number of doubly achieved distances two effect} (N22), respectively. The first statistic emphasizes triad relationships formed between actors with different covariate values, while $s_{i4}$ emphasizes indirect ties between actors. The covariate for the friendship data is the drinking behavior, and is the number of bills authored for the senate data. These additional effects are shown in Figure~\ref{fig:jtt} and Figure~\ref{fig:dad}, where the ties affected by the parameter value are represented by dotted lines.

@@

In this paper, we use three different model visualization techniques on SAOMs. First, Section~\ref{sec:minds} places the models in the data space. Then, Section~\ref{sec:collectmodels} explores visualizations of various collections of models. Finally, in Section~\ref{sec:algorithms} we use model visualization to explore the underlying algorithms in the SAOM fitting process. We conclude with a discussion in Section~\ref{sec:discussion}.
 %In terms of the senate data, then, the value of $\beta_3$ should dictate how willing a senator is to coauthor a bill on which the author of the bill has a different level of authorship, assuming there is an intermediary between the two who has the same authorship level as the first. %Comparison of the parameter estimates of these models fit to the two different data sets is given in Table~\ref{tab:meanests}.
%In addition, we fit two more models, M4 and M5 to the senate data that were found to be significant according to the Wald-type test:

@@

Every good data analysis includes both numerical and visual summaries of the data, so why restrict model description and diagnostics to numerical summaries? The concept of \emph{model visualization} was developed to complement traditional model diagnostic tools such as $R^2$ values and residual plots \citep{modelvis}. In a modeling task, there are three different things called ``model" under consideration: the model \emph{family}, the model \emph{form}, and the \emph{fitted} model. The latter is usally what first comes to mind when considering a model, where a specified model is fit, and parameter estimates are reported.% In the context of SAOMs, the \emph{fitted model} contains the form of the rate and objective functions, the estimated rate parameters, and the estimated objective function parameters. 
The \textit{model form} describes the the model \textit{before} the fitting process, defining which parameters are in the model within the context of the larger model family. % In SAOMs, the \emph{model form} includes description of the rate and objective functions and the variables therein that describe how the network evolves over time. 
Finally, the \textit{model family} is the broadest description of the model. This is the type of model that you wish to fit to the data, and is chosen based on the problem, data, and knowledge at hand. %For example, we chose to use a SAO \emph{model family} to model our network data over an exponential random graph model (ERGM) because we believe that actor-level variables effect network structure and formation, and we wanted to model the network changes over time.

The model family, the model form, and the fitted model can each be visualized according to the three principles of model visualization: we can view the model in the data space, visualize collections of models, and explore the process of fitting the model, not just the end result. Since we have already decided on our model family, SAOMs, we shift our focus to the fitted model and the model form. Specifically, we want to learn more about how the \emph{model form} we choose affects the \emph{fitted model} by using our example data sets and our visualization toolbox. %We begin by introducing the five models that we fit to our example data. 
We use the data and models described in Section~\ref{sec:exdatamods} to guide our visual explorations of SAOMs under the three principles of model visualization. %We begin with novel tools and ideas to view a SAOM in the data space of a dynamic network. We then explore collections of the same models fit many times to the example data to see how the simulation processes in the \texttt{RSiena} ``black-box" affect the model fits. Finally, we look behind the scenes and into the individual steps of the continuous time Markov chain to learn more about how this ``hidden data" mechanism works and how it results in a fitted model.

The first way we hope to better understand SAOMs is by viewing the model in the data space. In \citeauthor{modelvis}, they chose define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. But what does this definition mean for dynamic network models? We interpret the data space for our application as the combined data structures of the network: the node, edge, and time information.
%We want to make inferences on parameters in the model, but the multi-level data structure of networks makes the ``region of inference" difficult to define.
% %For dynamic social networks, there are a few different data ``spaces":
% \begin{enumerate}
% \item the actors and their corresponding covariates,
% \item the edges and their correpsonding variables that describe the relationships between the nodes, and
% \item the time, both the continuous unobserved time and the discrete observed time points, over which the network evolves
% \end{enumerate}
%The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if there is an edge between two nodes, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side.

@@

The node, edge, and time data can be visualized together in various ways, and one tool that can help is the \texttt{R} package \texttt{geomnet} \citep{geomnet}. The network's node or edge data can be mapped to different visual features in the node-link diagram.  The color, size, and shape of the points can be used to represent variables in the node data, while the color, linewidth, and linetype of the lines between points can be used to represent the edge variables. %In a social network, node data might be age, gender, and/or occupation of the person in the network, and edge data might be length of connection between two people, how the people first met (school, work, church, etc.), and/or how often they interact. 
To view temporal changes, we can view the network at different timepoints side-by-side to see the evolution. Pulling all of this information together with \texttt{geomnet} allows the entire data space to be viewed at once.

@@

To demonstrate, we use \texttt{geomnet} to visualize the connections in the $111^{th}$ United States Senate at two different points: when Hillary Clinton was in the senate, and after she left to become Secretary of State. Clinton was only in the $111^{th}$ senate from January 3-20, 2009, in the middle of her second term as the junior senator from New York. In that time, she authored two bills and was a cosponsor on 17s. There are many senators who cosponsored both of her bills, giving their edges to her a $WPC$ of 1. So with Clinton included in the node-link diagram in Figure~\ref{fig:funsenatevis}, the senate looks much more collaborative than it does without her. We then map our potential covariates to visual features: number of bills authored  maps to the size of the node, gender maps to the shape, and party to the color. In addition, we visualize the strength of the tie by mapping the $WPC$ value between the two senators to the linewidth of the edge. In this single visualization, we have viewed node information, edge information, and time.

@@

p1 <- ggplot(data = t1) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = "black",
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Average: M1") + ThemeNet

@@

%For network data, we view an ``average" network through a sort of summary network drawn using the traditional node-link diagram. 
In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1. To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then aggregate the 1,000 simulated instances of wave 2, and by counting up the number of times each edge $i \rightarrow j$ appears in a network simulation. Then, we construct a single network for each time point with edge weights equal to the proportion of time the edge appears in our 1,000 simulations. This weighted network is shown in the middle panel of Figure~\ref{fig:summnetM1}, with edges present only in the average network if it appears in at least 51 of the 1,000 simulations. The edges that appear most frequently in the model are drawn with darker, thicker lines. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data on the right of Figure~\ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

<<summary2, fig.height=2.5, fig.cap="Network differences between the average network from M1 and Wave 1 and Wave 2 of the data. Some edges appear in both the data and the average Wave 2 from M1, while others only appear in the data or only in the average. Width of edges in the average network represent the proportion of appearances in the 1,000 simulations. The FR layout is used.">>=

red <- "#e41a1c"
green <- "#4daf4a"

am_t1 <- as.matrix(igraph::as_adj(igraph::graph_from_data_frame(arrange(t1, parse_number(from), parse_number(to)))))[-17,-17]

diff1 <- am_t1 - fd2.w1
diff1[which(diff1==1)] <- 2
diff1[which(fd2.w1 == 1 & am_t1 == 1)]  <- 1
diff1 <- diff1 %>% dplyr::as_data_frame() 
names(diff1) <- paste0("V", 1:16)
diff1 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>%
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `2` = "InAvgNotW1",  `1`= "InBoth"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = green)) %>%
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_1 
t1_1 <- arrange(t1_1, from, to)
p1 <- ggplot(data = t1_1) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_1$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02, layout.alg = "fruchtermanreingold",
           singletons = T) +
  labs(title = "Difference between Wave 1 &\nWave 2 average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))

pforleg <- ggplot(data = mtcars) + 
  geom_line(aes(x = mpg, drat, group = as.factor(cyl), color = as.factor(cyl)), size = 1.5) + 
  scale_color_manual(values = c("black", red, green), name = "Tie is...", labels = c("in data,\nin average", "in data,\nnot in average", "not in data,\nin average")) + 
  theme_grey() + 
  theme(legend.key = element_rect(color = "black", fill = "white"), 
        legend.key.height = unit(25, "pt"))
#pforleg

diff2 <- -am_t1 + fd2.w2 
diff2[which(diff2==1)] <- 2
diff2[which(fd2.w2 == 1 & am_t1 == 1)]  <- 1
diff2 <- dplyr::as_data_frame(diff2)
names(diff2) <- paste0("V", 1:16)
diff2 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>% 
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `-1` = "InAvgNotW2",  `1`= "InBoth", `2` = "InW2notAvg"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = red, `-1` = green)) %>% 
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_2 
t1_2$linewidth[is.na(t1_2$linewidth) & !is.na(t1_2$to)] <- 1
t1_2 <- arrange(t1_2, from, to)
t1_2 <- t1_2 %>% filter(!(from == 16 & is.na(to)))
p2 <- ggplot(data = t1_2) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_2$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02, layout.alg = "fruchtermanreingold",
           singletons = T) +
  labs(title = "Difference between Wave 2 &\nWave 2 average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))
legendsumm <- cowplot::get_legend(pforleg)
ggdraw() + draw_plot(p1, width = .4) +  draw_plot(p2, width = .4, x=.4) + 
  draw_grob(legendsumm,x=.8, y=.25, width = .2, height = .4)
@

In Figure~\ref{fig:summary2}, we show \textit{difference networks} between the average network from M1 and the first and second waves of data. These networks are constructed by computing the difference between two adjacency matrices. The lack of model fit is very obvious here, with the green edges representing ties that only show up in the average network, and the red edges representing ties in the data that the model misses completely.  Again, the edgeweights demonstrate the frequency of the ties in the simulations from M1. On the left in Figure~\ref{fig:summary2}, we see the difference between the first wave of data and the average network. We only see green and black edges, meaning that the simulated networks consist primarily of edges in wave 1, plus some new edges that are not in the first wave of data. The additional green edges are all reciprocating black edges, meaning that the model predicts more reciprocity in ties than there actually is. On the right in Figure~\ref{fig:summary2}, we see the difference between the second wave of data and the average network. The model is simulating the second wave of data, so we would hope to see few differences between the average and wave2. However, there are six red edges, which are edges in wave 2 that do not appear at all in the average network, and there are several green edges that are in the average network, but not in the data. Although there are some edges that appear in both the average and wave 2, the majority of edges in the average are not in wave 2, and there are several missing edges. Thus, the model M1 is an incredibly poor fit, which we are able to see very easily by viewing the model in the data space in this way.

@@

The next method of viewing collections of models is with varying model settings. We have varied model settings already by choosing three different models M1, M2, M3 to fit to our example data sets. In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examined the parameter densities. Here, we simulate from these three models. For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model. We simulated 1,000 observations from each of the three models in this way. From these simulations, we created visualizations that represent an average network from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These averages are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave.

@@

\subsection*{Section 4.3}

\subsection*{Section 4.4}
